Central Limit Theorems for Super Ornstein-Uhlenbeck Processes with stable Branching

### 2019-07-30 supCLT.tex
change supCLT.tex

        Let the \emph{branching mechanism} $\psi$ be a function on $\mathbb R_+$ which takes the form of

@@

	We refer to $\pi$ as the \emph{L\'evy measure} of the branching mechanism $\psi$.

@@

    We say a real-valued Borel function $f:(t,x)\mapsto f(t,x)$ on $\mathbb R_+\times \mathbb R^d$ is \emph{locally bounded} if for each $t\in \mathbb R_+$, we have

@@

@@

       is a \emph{super Ornstein-Uhlenbeck process (super-OU process)} with branching mechanism $\psi$, if for each non-negative bounded Borel function $f$ on $\mathbb R^d$, we have

@@

   The strong law of large numbers for the super-OU process $X$ can be stated as follows:
    Under some conditions on the branching mechanism $\psi$ (these conditions are satisfied under our Assumptions 1 and 2 below), there exists an $\Omega_0$ of $\mathbb{P}_\mu$-full probability for every $\mu\in\mathcal M(\mathbb R^d)$ such that on $\Omega_0$,
    for every Lebesgue-almost everywhere continuous bounded non-negative function $f$ on $\mathbb R^d$,
    we have
    $\lim_{t\to\infty}e^{-\alpha t}\langle f, X_t\rangle=H_\infty\langle f, \varphi\rangle $,
    where $H_\infty$ is the limit of the non-negative martingale $e^{-\alpha t}\langle 1,X_t\rangle$ and
$\varphi$ is the invariant density of the OU process $\xi$ defined in \eqref{invariantdensity} below.
    See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.

    In this paper, we will establish some spatial central limit theorems (CLT) for the super-OU process $X$ above.
    Our key assumption is that the branching mechanism $\psi(z)$ takes the form of $-\alpha z +\rho z^2+ \psi_1(z)$, where $\psi_1$ is close to $\eta z^{1+\beta}$ with $\eta>0$ and $\beta\in (0, 1)$, in some sense.
    We want to find
    $(F_t)_{t\geq 0}$ and $(G_t)_{t\geq 0}$ such that
\[
    \frac{\langle f, X_t \rangle -G_t}{F_t}
\]
    converges weakly to some non-degenerate random variable as $t\rightarrow\infty$, for a large class of testing functions $f$.
    It turns out that the statements of the CLT are different in three regimes
    depending on the sign of $\alpha\beta-\kappa_f b (1+\beta)$: the small branching rate regime $\alpha\beta< \kappa_f b(1+\beta)$; the critical branching rate regime $\alpha\beta = \kappa_f b(1+\beta)$; and the large branching rate regime $\alpha\beta > \kappa_f b(1+\beta)$. Here, $\kappa_f$ is the order of $f$ in the spectral decomposition in terms of the spectrum of $L$.
    Note that, in the setting of this paper, $\langle f,X_t\rangle$ typically have infinite second moment.

    There are many papers studying central limit theorems for branching processes, branching diffusions and superprocesses under the second moment condition.
    See
    \cite{Heyde1970A-rate, HeydeBrown1871An-invariance, HeydeLeslie1971Improved} for supercritical Galton-Watson processes,
    \cite{KestenStigum1966Additional,KestenStigum1966A-limit} for supercritical multi-type Galton-Watson processes,
    \cite{Athreya1969Limit,Athreya1969LimitB,Athreya1971Some} for supercritical multi-type continuous time branching processes and \cite{AsmussenHering1983Branching} for general supercritical branching Markov processes under certain conditions.
    In \cite{AdamczakMilos2015CLT},
    Adamczak and Mi\l{}o\'s proved some spatial central limit theorems for 
    supercritical branching Ornstein-Uhlenbeck processes with binary branching mechanism and in \cite{Milos2012Spatial} 
    Mi\l{}o\'s proved some spatial central limit theorems for 
    supercritical super-OU processes with branching mechanisms satisfying a fourth moment condition.
    These two papers made connections between the CLT and  the branching rate regimes.
    In \cite{RenSongZhang2014Central}, Ren, Song and Zhang proved some spatial central limit theorems for supercritical super-OU 
    processes with branching mechanisms satisfying 
    only a second moment condition.
    Moreover,
    compared with the results of \cite{AdamczakMilos2015CLT,Milos2012Spatial}, the limit distributions in \cite{RenSongZhang2014Central} are non-degenerate.
    In \cite{RenSongZhang2014CentralB,RenSongZhang2015Central,RenSongZhang2017Central},
    Ren, Song and Zhang also established a series of spatial central limit theorems for a large class of general 
    supercritical branching Markov processes and superprocesses with spatially dependent branching mechanisms.
    The functional version of the central limit theorems was established in \cite{Janson2004Functional} for supercritical multitype branching processes, and  in \cite{RenSongZhang2017Functional} for supercritical superprocesses.

    There are also many limiting theorem type results 
    for supercritical branching processes and branching Markov processes
    with branching mechanisms of infinite second moments.
    Heyde \cite{Heyde1971Some} established a central limit type  theorem 
    for supercritical Galton-Watson process when the offspring distribution
    belongs to the domain of attraction of a stable law of index $\alpha\in (1, 2]$,
    and   proved that the limit law is  a stable law. Similar results  for supercritical multi-type Galton-Watson processes and supercritical  continuous time branching processes, under some $p$-th ($p\in(1,2]$) moment condition on the offspring distribution,
    were given in Asmussen \cite{Asmussen76Convergence}.
    Recently,   Marks and Milo\'s \cite{MarksMilos2018CLT}
    considered the limit behavior of supercritical branching Ornstein-Uhlenbeck processes with a special stable branching law.
	They established some spatial central limit theorems in the small and critical branching rate regimes,
    but they did not prove any central limit theorem type result in the large branching rate regime.
    We also mention here that very recently \cite{IksanovKoleskoMeiners2018Stable-like} considered stable fluctuations of Biggins' martingales in the context of branching random walks and \cite{RenSongSun2018Limit} considered the asymptotic behavior of a class of critical superprocesses with spatially dependent stable branching.

@@

We will always assume that the following assumption holds.
\begin{asp}
\label{asp: Greys condition}
	The branching mechanism satisfies Grey's condition, i.e., there exists $z' > 0$ such that $\psi(z) > 0$ for all $z>z'$ and  $\int_{z'}^\infty \psi(z)^{-1}dz < \infty$.
\end{asp}
    For each $\mu \in \mathcal M(\mathbb R^d)$, write $\|\mu\| = \langle 1, \mu\rangle$.
	It is known (see \cite[Theorems 12.5 \& 12.7]{Kyprianou2014Fluctuations} for example) that, under Assumption \ref{asp: Greys condition},
	the \emph{extinction event}
	$D:=\{\exists t\geq 0,~\text{s.t.}~\|X_t\|=0\}$ has positive probability, with respect to $\mathbb P_\mu$ for each  $\mu \in \mathcal M(\mathbb R^d)$.
    In fact,
\begin{equation}
    \mathbb{P}_{\mu} (D)
    = e^{-\bar v \|\mu\|},
    \quad \mu\in \mathcal M(\mathbb R^d),
\end{equation}
    where
\[
	\bar v := \sup\{\lambda \geq 0: \psi(\lambda) = 0\} \in (0,\infty)
\]
	is the largest root of $\psi$.

@@

    Assumption \ref{asp: branching mechanism} says that
 there exist constants $\eta>0$ and $\beta > 0$
    such that the L\'evy measure $\pi(dy)$ is not too far away from the measure $\eta \Gamma(-1-\beta)^{-1}y^{-2-\beta} dy$.
    In particular, if $\pi(dy)$ is equal to $\eta \Gamma(-1-\beta)^{-1}y^{-2-\beta} dy$, then the branching mechanism $\psi(z)$
    takes the form of
	$-\alpha z + \rho z^2 + \eta z^{1+\beta}$.
	It will be proved in Lemma \ref{lem: LlogL criterion} that, under Assumption
	\ref{asp: branching mechanism}, the branching mechanism $\psi$ satisfies the
	$L \log L$ condition, i.e.,
\[
	\int_{(1,\infty)} y\log y~\pi(dy) < \infty.
\]
    This guarantees that $H_\infty$, the limit of the non-negative martingale $(e^{-\alpha t} \|X_t\|)_{t\geq 0}$, is non-degenerate.

@@

    We now recall some basic spectral properties of $(P_t)_{t\geq 0}$ from \cite{MetafunePallaraPriola2002Spectrum}.
       It is known that the OU process $\xi$ has an invariant density
\begin{equation}
\label{invariantdensity}
    \varphi(x)dx
    :=\Big (\frac{b}{\pi \sigma^2}\Big )^{d/2}\exp \Big(-\frac{b}{\sigma^2}|x|^2 \Big)dx,
    \quad x\in \mathbb R^d.
\end{equation}
    Let $L^2(\varphi):= \left\{ h  \in \mathcal B(\mathbb R^d, \mathbb R): \int_{\mathbb R^d} |h(x)|^2 \varphi(x) dx < \infty \right\}$.
        Then $L^2(\varphi)$ is a Hilbert space with inner product
\begin{equation}
    \langle f_1, f_2 \rangle_{\varphi}
    := \int_{\mathbb R^d}f_1(x)f_2(x)\varphi(x) dx, \quad f_1,f_2 \in L^2(\varphi).
\end{equation}
       Let $\mathbb Z_+ := \mathbb N\cup\{0\}$.
    For each $p = (p_k)_{k = 1}^d \in \mathbb{Z}_+^{d}$, write $|p|:=\sum_{k=1}^d p_k$,
    $p!:= \prod_{k= 1}^d p_k!$ and $\partial x^p:= \prod_{k = 1}^d\partial x_k^{p_k}$.
    The \emph{Hermite polynomials} are defined by
\begin{equation}
    H_p(x)
    :=(-1)^{|p|}\exp(|x|^2) \frac{\partial ^{|p|}}{\partial x^p} \exp(-|x|^2) ,
    \quad x\in \mathbb R^d,
    p \in \mathbb{Z}_+^{d}.
\end{equation}
It is known that $(P_t)_{t\geq 0}$
       is a strongly continuous semigroup in $L^2(\varphi)$ and its generator $L$ has discrete spectrum
        $\sigma(L)= \{-bk: k \in \mathbb Z_+\}$.
    For each $k \in \mathbb Z_+$, denote by $\mathcal{A}_k$ the eigenspace corresponding to the eigenvalue $-bk$, then
$
    \mathcal{A}_k
    = \operatorname{Span} \{\phi_p : p\in \mathbb Z_+^d, |p|=k\},
$
    where
\begin{equation}\label{eigenfunction}
    \phi_p(x)
        := \frac{1}{\sqrt{ p! 2^{|p|} }} H_p \Big(\frac{ \sqrt{b} }{\sigma}x \Big),
    \quad x\in \mathbb R^d, p\in \mathbb Z_+^d.
\end{equation}
       In other words,
\begin{equation}
    P_t\phi_p(x)
    =e^{-b|p|t}\phi_p(x),
    \quad t\geq 0, x\in \mathbb R^d, p\in \mathbb Z_+^d.
\end{equation}
    Moreover, $\{\phi_p: p \in \mathbb Z_+^d\}$ forms a complete orthonormal basis of $L^2(\varphi)$.
    Thus for each $f\in L^2(\varphi)$, we have
\begin{equation}\label{semicomp1}
    f
         =\sum_{k=0}^{\infty}\sum_{p\in \mathbb Z_+^d:|p|=k}\langle f, \phi_p \rangle_{\varphi} \phi_p,
    \quad \text{in~} L^2(\varphi).
\end{equation}
    For each function $f\in L^2(\varphi)$, denote by
\begin{equation}
    \kappa_f
    :=\inf \left \{k\geq 0: \exists ~ p\in \mathbb Z_+^d ,{\rm ~s.t.~}|p|=k {\rm ~and~}  \langle f, \phi_p \rangle_{\varphi}\neq 0\right \},
\end{equation}
    the order of the function $f$ in the spectral decomposition \eqref{semicomp1}.
    Note that $ \kappa_f\geq 0$ and that, if $f\in L^2(\varphi)$ is non-trivial, then $\kappa_f<\infty$.
    In particular, the order of any constant non-zero function is zero.
    Denote by $\mathcal P$ the class of functions of polynomial growth on $\mathbb R^d$, i.e.,
\begin{equation}
\label{eq: polynomial growth function}
    \mathcal{P}
    :=\big \{f\in \mathscr B(\mathbb R^d, \mathbb R):\exists~ C>0, n \in \mathbb Z_+, {\rm such  \, \, that \, } \forall x\in \mathbb R^d,~ |f(x)|\leq C(1+|x|)^n\big\}.
\end{equation}
    It is clear that $\mathcal{P} \subset L^2(\varphi)$.
    In this paper, we are mainly interested in the asymptotic behavior of $\langle f, X_t\rangle$
    for non-trivial $f\in \mathcal P$.
    The asymptotic behavior varies with the following three regimes:
    \begin{itemize}
\item
    The large branching rate regime: $\alpha\beta>\kappa_fb(1+\beta)$;
\item
    The critical branching rate regime: $\alpha\beta=\kappa_fb(1+\beta)$;
\item
    The small branching rate regime: $\alpha\beta<\kappa_fb(1+\beta)$.
\end{itemize}

\subsubsection{Large branching rate regime}
    Denote by $\mathcal M_c(\mathbb R^d)$ the space of all finite Borel measures of compact support on $\mathbb R^d$.
    For each $p\in \mathbb{Z}_+^d$, define
\[
    H_t^p
    := e^{-(\alpha-|p|b)t}\langle\phi_p,X_t\rangle,
    \quad t\geq 0.
\]
    If $\alpha\beta>|p|b(1+\beta)$, then for all $\gamma\in (0, \beta)$ and $\mu\in \mathcal M_c(\mathbb R^d)$,
      we will prove in Lemma \ref{lemma26} that $(H_t^p)_{t\geq 0}$ is a $\mathbb{P}_{\mu}$-martingale bounded in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
    Thus the limit $H^p_{\infty}:=\lim_{t\rightarrow \infty}H_t^p$ exists $\mathbb{P}_{\mu}$-almost surely and in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
 \begin{thm}
 \label{thm: law of large number}
     If $f \in \mathcal{P}$ satisfies $\alpha\beta>\kappa_fb(1+\beta)$, then
   for all
     $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$,
\[
    e^{-(\alpha-\kappa_fb)t}\langle f, X_t\rangle
       \xrightarrow[t\to \infty]{}\sum_{p\in \mathbb Z_+^d:|p|=\kappa_f}\langle f, \phi_p\rangle_{\varphi} H_{\infty}^p
    \quad in~ L^{1+\gamma}(\mathbb{P}_{\mu}).
\]
    Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have almost sure convergence.
\end{thm}
\begin{rem}
   If $f\in \mathcal B(\mathbb R^d, \mathbb R_+)$ is non-trivial and  bounded, then $\kappa_f=0$.
    Hence, Theorem \ref{thm: law of large number} says that
    for any $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$, as $t\rightarrow \infty$,
\[
    e^{-\alpha t}\langle f, X_t\rangle
    \rightarrow \langle f, \varphi\rangle H_{\infty}
    \quad \mbox{in } L^{1+\gamma}(\mathbb{P}_{\mu}).
\]
    Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have almost sure convergence.
    However, to get  almost sure convergence for bounded, non-negative, and Lebesgue-almost everywhere continuous testing functions $f$, we do not need $f$ to be twice differentiable. See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.
\end{rem}
\begin{rem}
    For general test functions $f\in \mathcal{P}$ with $\alpha\beta>\kappa_fb(1+\beta)$, Theorem \ref{thm: law of large number} extends the strong laws of large numbers of \cite{ChenRenYang2019Skeleton} and \cite{EckhoffKyprianouWinkel2015Spines} in which the first order asymptotic ($\kappa_f=0$) was identified.
    \end{rem}

>>>>

%        Let the \emph{branching mechanism} $\psi$ be a function on $\mathbb R_+$ which takes the form of
 Let $\psi$ be a function on $\mathbb R_+$ of the form 

@@

%	We refer to $\pi$ as the \emph{L\'evy measure} of the branching mechanism $\psi$.
$\psi$ is referred to as a branching mechanism and $\pi$ is referred to as the L\'evy measure
of $\psi$.

@@

    We say a real-valued Borel function $f:(t,x)\mapsto f(t,x)$ on $\mathbb R_+\times \mathbb R^d$ is \emph{locally bounded} if, for each $t\in \mathbb R_+$, we have
%\[
$

@@

%\]
$

@@

       is a \emph{super Ornstein-Uhlenbeck process (super-OU process)} with branching mechanism $\psi$, 
     or a $(\xi, \psi)$ superprocess,  
       if for each non-negative bounded Borel function $f$ on $\mathbb R^d$, we have

@@

   The strong law of large numbers for the super-OU process $X$ can be stated as follows:
    Under some conditions on the branching mechanism $\psi$ (these conditions are satisfied under our Assumptions 1 and 2 below), there exists an $\Omega_0$ of $\mathbb{P}_\mu$-full probability for every $\mu\in\mathcal M(\mathbb R^d)$ such that on $\Omega_0$,
    for every Lebesgue-a.\/e. continuous bounded non-negative function $f$ on $\mathbb R^d$,
    we have
    $\lim_{t\to\infty}e^{-\alpha t}\langle f, X_t\rangle=H_\infty\langle f, \varphi\rangle $,
    where $H_\infty$ is the limit of the martingale $e^{-\alpha t}\langle 1,X_t\rangle$ and
$\varphi$ is the invariant density of the OU process $\xi$ defined in \eqref{invariantdensity} below.
    See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.

    In this paper, we will establish some spatial central limit theorems (CLT) for the super-OU process $X$ above.
    Our key assumption is that the branching mechanism $\psi(z)$ takes the form of $-\alpha z +\rho z^2+ \psi_1(z)$, where $\psi_1$ is close to $\eta z^{1+\beta}$ with $\eta>0$ and $\beta\in (0, 1)$, in some sense.
    We want to find
    $(F_t)_{t\geq 0}$ and $(G_t)_{t\geq 0}$ such that
%\[
$
 %   \frac{\langle f, X_t \rangle -G_t}{F_t}
    (\langle f, X_t \rangle -G_t)/F_t
%\]
$
    converges weakly to some non-degenerate random variable as $t\rightarrow\infty$, for a large class of functions $f$.
    It turns out that the statements of the CLT's are different in three regimes
    depending on the sign of $\alpha\beta-\kappa_f b (1+\beta)$: the small branching rate regime $\alpha\beta< \kappa_f b(1+\beta)$; the critical branching rate regime $\alpha\beta = \kappa_f b(1+\beta)$; and the large branching rate regime $\alpha\beta > \kappa_f b(1+\beta)$. Here, $\kappa_f$ is the order of $f$ in the spectral decomposition in terms of the spectrum of $L$.
    Note that, in the setting of this paper, $\langle f,X_t\rangle$ typically have infinite second moment.

    There are many papers studying central limit theorems for branching processes, branching diffusions and superprocesses under the second moment condition.
   See   \cite{Heyde1970A-rate, HeydeBrown1871An-invariance, HeydeLeslie1971Improved} for %supercritical Galton-Watson processes,
   supercritical Galton-Watson processes (GW processes),
    \cite{KestenStigum1966Additional,KestenStigum1966A-limit} for supercritical multi-type %Galton-Watson processes,
    GW processes
    \cite{Athreya1969Limit,Athreya1969LimitB,Athreya1971Some} for supercritical multi-type continuous time branching processes and \cite{AsmussenHering1983Branching} for general supercritical branching Markov processes under certain conditions.
%In \cite{AdamczakMilos2015CLT}, Adamczak and Mi\l{}o\'s proved some spatial central limit theorems for   supercritical branching Ornstein-Uhlenbeck processes with binary branching mechanism and in \cite{Milos2012Spatial} Mi\l{}o\'s proved some spatial central limit theorems for supercritical super-OU processes with branching mechanisms satisfying a fourth moment condition.
Some spatial central limit theorems for   supercritical branching OU processes with binary branching mechanism were proved in \cite{AdamczakMilos2015CLT} and some spatial central limit theorems for supercritical super-OU processes with branching mechanisms satisfying a fourth moment condition were proved in \cite{Milos2012Spatial}.
    These two papers made connections between the CLT and  the branching rate regimes.
%In \cite{RenSongZhang2014Central}, Ren, Song and Zhang proved some spatial central limit theorems for supercritical super-OU  processes with branching mechanisms satisfying only asecond moment condition.
Some spatial central limit theorems for supercritical super-OU  processes with branching mechanisms satisfying only asecond moment condition were established in \cite{RenSongZhang2014Central}.
Moreover, compared with the results of \cite{AdamczakMilos2015CLT,Milos2012Spatial}, the limit distributions in \cite{RenSongZhang2014Central} are non-degenerate.
%In \cite{RenSongZhang2014CentralB,RenSongZhang2015Central,RenSongZhang2017Central}, Ren, Song and Zhang also established a series of spatial central limit theorems for a large class of general supercritical branching Markov processes and superprocesses with spatially dependent branching mechanisms.
Since then, a series of spatial central limit theorems for a large class of general supercritical branching Markov processes and superprocesses with spatially dependent branching mechanisms
were proved in \cite{RenSongZhang2014CentralB,RenSongZhang2015Central,RenSongZhang2017Central}.
    The functional version of the central limit theorems was established in \cite{Janson2004Functional} for supercritical multitype branching processes, and  in \cite{RenSongZhang2017Functional} for supercritical superprocesses.

There are also many limiting theorem type results for supercritical branching processes and branching Markov processes with branching mechanisms of infinite second moments.
Heyde \cite{Heyde1971Some} established a central limit type  theorem 
%for supercritical Galton-Watson process when the offspring distribution
for supercritical GW processes when the offspring distribution
belongs to the domain of attraction of a stable law of index $\alpha\in (1, 2]$,
and   proved that the limit law is  a stable law. Similar results  for supercritical multi-type 
%Galton-Watson processes and supercritical  continuous time branching processes, 
GW processes and supercritical  continuous time branching processes, 
under some $p$-th ($p\in(1,2]$) moment condition on the offspring distribution,
were given in Asmussen \cite{Asmussen76Convergence}.
Recently,   Marks and Milo\'s \cite{MarksMilos2018CLT}
%considered the limit behavior of supercritical branching Ornstein-Uhlenbeck processes 
considered the limit behavior of supercritical branching OU processes
with a special stable branching law.
	They established some spatial central limit theorems in the small and critical branching rate regimes,
    but they did not prove any central limit theorem type result in the large branching rate regime.
    We also mention here that very recently \cite{IksanovKoleskoMeiners2018Stable-like} considered stable fluctuations of Biggins' martingales in the context of branching random walks and \cite{RenSongSun2018Limit} considered the asymptotic behavior of a class of critical superprocesses with spatially dependent stable branching.

@@

We will always assume that the following assumption holds.
\begin{asp}
\label{asp: Greys condition}
	The branching mechanism satisfies Grey's condition, i.e., there exists $z' > 0$ such that $\psi(z) > 0$ for all $z>z'$ and  $\int_{z'}^\infty \psi(z)^{-1}dz < \infty$.
\end{asp}
    For each $\mu \in \mathcal M(\mathbb R^d)$, write $\|\mu\| = \langle 1, \mu\rangle$.
	It is known (see \cite[Theorems 12.5 \& 12.7]{Kyprianou2014Fluctuations} for example) that, under Assumption \ref{asp: Greys condition},
	the \emph{extinction event}
	$D:=\{\exists t\geq 0,~\text{s.t.}~\|X_t\|=0\}$ has positive probability, with respect to $\mathbb P_\mu$ for each  $\mu \in \mathcal M(\mathbb R^d)$.
    In fact,
%\begin{equation}
$
    \mathbb{P}_{\mu} (D)
    = e^{-\bar v \|\mu\|},
%    \quad \mu\in \mathcal M(\mathbb R^d),
%\end{equation}
$ for $\mu\in \mathcal M(\mathbb R^d)$,
    where
%\[
$
	\bar v := \sup\{\lambda \geq 0: \psi(\lambda) = 0\} \in (0,\infty)
%\]
$
	is the largest root of $\psi$.

@@

    Assumption \ref{asp: branching mechanism} says that
 there exist constants $\eta>0$ and $\beta > 0$
    such that the L\'evy measure $\pi(dy)$ is not too far away from the measure $\eta \Gamma(-1-\beta)^{-1}y^{-2-\beta} dy$.
    In particular, if $\pi(dy)$ is equal to $\eta \Gamma(-1-\beta)^{-1}y^{-2-\beta} dy$, then the branching mechanism $\psi(z)$
    takes the form of
	$-\alpha z + \rho z^2 + \eta z^{1+\beta}$.
	It will be proved in Lemma \ref{lem: LlogL criterion} that, under Assumption
	\ref{asp: branching mechanism}, the branching mechanism $\psi$ satisfies the
	$L \log L$ condition, i.e.,
%\[
$
	\int_{(1,\infty)} y\log y~\pi(dy) < \infty.
%\]
$
    This guarantees that $H_\infty$, the limit of the non-negative martingale $(e^{-\alpha t} \|X_t\|)_{t\geq 0}$, is non-degenerate.

@@

    We now recall some basic spectral properties of $(P_t)_{t\geq 0}$ from \cite{MetafunePallaraPriola2002Spectrum}.
       It is known that the OU process $\xi$ has an invariant density
\label{invariantdensity}
    \varphi(x)dx
    :=\Big (\frac{b}{\pi \sigma^2}\Big )^{d/2}\exp \Big(-\frac{b}{\sigma^2}|x|^2 \Big)dx,
    Let $L^2(\varphi):= \left\{ h  \in \mathcal B(\mathbb R^d, \mathbb R): \int_{\mathbb R^d} |h(x)|^2 \varphi(x) dx < \infty \right\}$.
        Then $L^2(\varphi)$ is a Hilbert space with inner product
    \langle f_1, f_2 \rangle_{\varphi}
    := \int_{\mathbb R^d}f_1(x)f_2(x)\varphi(x) dx, \quad f_1,f_2 \in L^2(\varphi).
       Let $\mathbb Z_+ := \mathbb N\cup\{0\}$.
    For each $p = (p_k)_{k = 1}^d \in \mathbb{Z}_+^{d}$, write $|p|:=\sum_{k=1}^d p_k$,
    $p!:= \prod_{k= 1}^d p_k!$ and $\partial x^p:= \prod_{k = 1}^d\partial x_k^{p_k}$.
    The \emph{Hermite polynomials} are defined by
    H_p(x)
    :=(-1)^{|p|}\exp(|x|^2) \frac{\partial ^{|p|}}{\partial x^p} \exp(-|x|^2) ,
    \quad x\in \mathbb R^d,
    p \in \mathbb{Z}_+^{d}.
It is known that $(P_t)_{t\geq 0}$
       is a strongly continuous semigroup in $L^2(\varphi)$ and its generator $L$ has discrete spectrum
        $\sigma(L)= \{-bk: k \in \mathbb Z_+\}$.
    For each $k \in \mathbb Z_+$, denote by $\mathcal{A}_k$ the eigenspace corresponding to the eigenvalue $-bk$, then
$
    \mathcal{A}_k
    = \operatorname{Span} \{\phi_p : p\in \mathbb Z_+^d, |p|=k\},
$
    where
\begin{equation}\label{eigenfunction}
    \phi_p(x)
        := \frac{1}{\sqrt{ p! 2^{|p|} }} H_p \Big(\frac{ \sqrt{b} }{\sigma}x \Big),
    \quad x\in \mathbb R^d, p\in \mathbb Z_+^d.
       In other words,
    P_t\phi_p(x)
    =e^{-b|p|t}\phi_p(x),
    \quad t\geq 0, x\in \mathbb R^d, p\in \mathbb Z_+^d.
    Moreover, $\{\phi_p: p \in \mathbb Z_+^d\}$ forms a complete orthonormal basis of $L^2(\varphi)$.
    Thus for each $f\in L^2(\varphi)$, we have
\begin{equation}\label{semicomp1}
    f
         =\sum_{k=0}^{\infty}\sum_{p\in \mathbb Z_+^d:|p|=k}\langle f, \phi_p \rangle_{\varphi} \phi_p,
    \quad \text{in~} L^2(\varphi).
    For each function $f\in L^2(\varphi)$, denote by
    \kappa_f
    :=\inf \left \{k\geq 0: \exists ~ p\in \mathbb Z_+^d ,{\rm ~s.t.~}|p|=k {\rm ~and~}  \langle f, \phi_p \rangle_{\varphi}\neq 0\right \},
    the order of the function $f$ in the spectral decomposition \eqref{semicomp1}.
%new
$\kappa_f$ corresponds to the lowest non-trivial frequency in the eigen-expansion of $f$.
%end new    
    Note that $ \kappa_f\geq 0$ and that, if $f\in L^2(\varphi)$ is non-trivial, then $\kappa_f<\infty$.
    In particular, the order of any constant non-zero function is zero.
    Denote by $\mathcal P$ the class of functions of polynomial growth on $\mathbb R^d$, i.e.,
\label{eq: polynomial growth function}
    \mathcal{P}
    :=\big \{f\in \mathscr B(\mathbb R^d, \mathbb R):\exists~ C>0, n \in \mathbb Z_+, {\rm such  \, \, that \, } \forall x\in \mathbb R^d,~ |f(x)|\leq C(1+|x|)^n\big\}.
    It is clear that $\mathcal{P} \subset L^2(\varphi)$.
    In this paper, we are mainly interested in the asymptotic behavior of $\langle f, X_t\rangle$
    for non-trivial $f\in \mathcal P$.
    The asymptotic behavior varies with the following three regimes:
%    \begin{itemize}
%\item
%    The large branching rate regime: $\alpha\beta>\kappa_fb(1+\beta)$;
%\item
%    The critical branching rate regime: $\alpha\beta=\kappa_fb(1+\beta)$;
%\item
%    The small branching rate regime: $\alpha\beta<\kappa_fb(1+\beta)$.
%\end{itemize}
(1) the small branching rate regime where $\alpha\beta<\kappa_fb(1+\beta)$; (2) the 
critical branching rate regime where $\alpha\beta=\kappa_fb(1+\beta)$; and
(3) the large branching rate regime where $\alpha\beta>\kappa_fb(1+\beta)$.

\subsubsection{Large branching rate regime}
    Denote by $\mathcal M_c(\mathbb R^d)$ the space of all finite Borel measures of compact support on $\mathbb R^d$.
    For each $p\in \mathbb{Z}_+^d$, define
\[
    H_t^p
    := e^{-(\alpha-|p|b)t}\langle\phi_p,X_t\rangle,
    \quad t\geq 0.
\]
    If $\alpha\beta>|p|b(1+\beta)$, then for all $\gamma\in (0, \beta)$ and $\mu\in \mathcal M_c(\mathbb R^d)$,
      we will prove in Lemma \ref{lemma26} that $(H_t^p)_{t\geq 0}$ is a $\mathbb{P}_{\mu}$-martingale bounded in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
    Thus the limit $H^p_{\infty}:=\lim_{t\rightarrow \infty}H_t^p$ exists $\mathbb{P}_{\mu}$-almost surely and in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
 \begin{thm}
 \label{thm: law of large number}
     If $f \in \mathcal{P}$ satisfies $\alpha\beta>\kappa_fb(1+\beta)$, then
   for all
     $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$,
\[
    e^{-(\alpha-\kappa_fb)t}\langle f, X_t\rangle
       \xrightarrow[t\to \infty]{}\sum_{p\in \mathbb Z_+^d:|p|=\kappa_f}\langle f, \phi_p\rangle_{\varphi} H_{\infty}^p
    \quad in~ L^{1+\gamma}(\mathbb{P}_{\mu}).
\]
    Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have almost sure convergence.
\end{thm}
\begin{rem}
   If $f\in \mathcal B(\mathbb R^d, \mathbb R_+)$ is non-trivial and  bounded, then $\kappa_f=0$.
    Hence, Theorem \ref{thm: law of large number} says that
    for any $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$, as $t\rightarrow \infty$,
\[
    e^{-\alpha t}\langle f, X_t\rangle
    \rightarrow \langle f, \varphi\rangle H_{\infty}
    \quad \mbox{in } L^{1+\gamma}(\mathbb{P}_{\mu}).
\]
    Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have a.\/s.\ convergence.
    However, to get  a.\/s.\ convergence for bounded non-negative 
    Lebesgue-a.\/e.\/\ continuous functions $f$, we do not need $f$ to be twice differentiable. See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.
\end{rem}
\begin{rem}
    For general $f\in \mathcal{P}$ with $\alpha\beta>\kappa_fb(1+\beta)$, Theorem \ref{thm: law of large number} extends the strong laws of large numbers of \cite{ChenRenYang2019Skeleton, EckhoffKyprianouWinkel2015Spines} in which 
the first order asymptotic ($\kappa_f=0$) was identified.
    \end{rem}
